Switchgrass Phenology and Genotype-by-Environment Mapping

### 2023-12-28 manuscript/revision2/revision2.qmd
manuscript: Edit Abstract and Significance statement

        postal-code: 

@@

        postal-code: 

@@

        postal-code: 

@@

        postal-code: 

@@

        postal-code: 

@@

        postal-code: 

@@

        postal-code: 

@@

  The timing of vegetative growth and flowering in plants ("phenological timings") multiple depend on environmental cues; thus, phenological traits can have important genotype-by-environment (GxE) interactions. Here, we map genotype-by-weather effects on phenological timings in multiple populations of two highly divergent switchgrass (Panicum virgatum) populations spanning the central United States. We distinguish GxWeather patterns covarying with interpretable weather-based cues and agnostic, empirically-inferred GxWeather patterns. Most GxWeather effects belong to the latter category and do not covary with weather-based cues. However, >XX% of Gulf population effects on flowering time covary with photoperiod cues, while >XX% of Midwest upland population SNPs affecting flowering had effects that covaried with XX cues. An independent pseudo-F2 cross of Gulf and Midwest individuals mapped 23 additive QTLs for flowering at the same common gardens, all with significant mash associations and ten with enrichment of highly significant mash associations. We demonstrate that we can identify QTL with GxE and assign them to specific weather-based cues and/or data-driven patterns. Breeding for particular alleles at these loci could change flowering responsiveness to photoperiod cues in switchgrass. More broadly, this approach could be used to identify genetic marker-environment interactions in any species with related populations phenotyped in multiple environments.

@@

  The timing of plant seasonal development (phenology) has major impacts on fitness because of the steep price of plant-environment mismatches. Here we map the genetic basis of two phenological events, the start of above-ground growth and flowering, in two genetically and phenologically distinct populations of switchgrass. We do this at eight field locations spanning the latitudinal range of both switchgrass populations. Our approach allows us to identify regions of the genome that covary with specific weather-related environmental features at every location. For flowering, these features differed by population: the Midwest population had genetic variation that primarily covaried with cumulative growing degree days, a temperature-related measure, while the Gulf population had genetic variation that primarily covaried with photoperiod, a day-length-related measure.

@@

keywords:  [allele-by-environment effect variation, antagonistic pleiotropy, photoperiod, cumulative growing degree days, genetic variation]

@@

\firstpage[15]{2}

@@

Flowering time, in particular, is a common subject of GxE research [@blackman2013interacting; @henry2014transitions; @aagren2017adaptive; @brachi2010linkage; @dittmar2014flowering; @li2018genomic; @romero2017study], a key output of selection driving adaptation to local environments [@andres2012genetic; @wadgymar2017identifying; @blumel2015flowering], and a selection target for crop improvement to adapt crops to local or future environments @jung2009flowering. Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in a number of cereal species [@turner2005pseudo; @faure2012mutation; @hung2012zmcct; @zakhrabekova2012induced; @yang2013oself3] and other crops [@pin2012multifaceted; @weller2019parallel]. Recent statistical advances in studying phenological GxE have involved determining critical environmental indices before the phenological event occurs, such as photothermal time within a critical growth window @li2018genomic. However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE. Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying], but has been limited by a known statistical bias that reduced detection of SNP effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary]. Thus, despite substantial interest in the frequencies of various forms of GxE, the prevalence of antagonistic pleiotropy relative to other forms of GxE remains unknown.

@@

In our diversity panel of tetraploid switchgrass [@lovell2021genomic], genotypes from the Gulf and Midwest genetic subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (@fig-map). At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, Gulf green-up occurred before Midwestern green-up, and Gulf flowering occurred after Midwestern flowering (Fig. 1A). At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both Gulf green-up and flowering occurred after Midwest green-up and flowering. At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up occurred over the same time period. These patterns led to strong negative phenotypic correlations for green-up between the North and Texas gardens, particularly in the Gulf and Both subpopulations, and contributed to positive phenotypic correlations for flowering time of larger magnitude at more northern gardens (Fig. 1B).

Narrow-sense heritabilities (h2) indicated that rank-changing GxE for these phenotypes was present across the common gardens (Fig 1C). h2 were typically high at individual gardens: 59% on average for green-up date, and 87% for flowering date. However, h2 were variable across gardens, and green-up dates were uncorrelated (r2 \< 0.2) or negatively correlated between pairs of gardens (Fig. 1B). These negative and small correlations undoubtedly contributed to the low h2 values for green-up and flowering date when estimated jointly at all eight gardens: h2 was 0.8% for green-up and 23.2% for flowering date.

@@

We next explored how genetic variation in phenology covaried with environmental cues by using mash to jointly discover trait-associated SNPs across all common gardens for the Midwest, Gulf, and 'Both' subpopulations (Fig 1D, SI Appendix, Datasets 1-6). We first estimated SNP effects using GWAS on site-specific BLUPs for each phenotype. We then used mash on two subsets of these SNP effects, first to determine the covariance structures that significantly improved the mash model log likelihood when included, then to re-estimate the set of SNP effects with the largest effect estimates in univariate GWAS (hereafter "strong" effects) using the set of significant covariance matrices.

>>>>

        postal-code: 57006

@@

        postal-code: 76502

@@

        postal-code: 65211

@@

        postal-code: 48824

@@

        postal-code: 68583

@@

        postal-code: 75684

@@

        postal-code: 74078

@@

  The timing of vegetative growth and flowering in plants ("phenological timings") depend on both genetic variation and environmental cues; thus, phenological traits can have important genotype-by-environment (GxE) interactions. In particular, understanding the mixture and prevalence of genotype-by-weather (GxWeather) interactions primed by weather prior to the phenological event should aid prediction and manipulation of phenological timings. Here, we map genotype-by-weather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of these populations at sites spanning the central United States. We distinguish GxWeather patterns covarying with interpretable weather-based cues from agnostic, site-based patterns. Most GxWeather effects belong to the latter category and do not covary with weather-based cues. However, 65% of Gulf population effects on the start of vegetative growth covary with daylength 14 days prior to green-up date. 33% of Gulf population effects on flowering date covary with cumulative rainfall in the seven days prior to flowering, while 22% of Midwest population effects on flowering date covaried with day length change two days prior to flowering. An independent pseudo-F2 cross of Gulf and Midwest individuals mapped 23 additive QTLs for flowering at the same common gardens, all with significant mash associations and eleven with enrichment of highly significant mash associations. We demonstrate that we can identify QTL with GxWeather and assign them to specific weather-based cues or other patterns. Breeding for particular alleles at these loci could change flowering responsiveness to photoperiod cues in switchgrass. More broadly, this approach could be used to identify genetic marker-environment interactions in any species with related populations phenotyped in multiple environments.

@@

  The timing of plant seasonal development (phenology) has major impacts on fitness because of the steep price of plant-environment mismatches. We infer the mixture of ways genetic effects on phenological traits depend on plant environment, particularly how effects covary with weather prior to the phenological event (GxWeather). Most effects do not have GxWeather, but a minority do. GxWeather is population-specific. The majority of effects on the start of vegetative growth in the Gulf population have GxWeather: these effects differ in sign across environments and covary with a photoperiod cue 14 days prior. A minority of effects on flowering date in the Gulf and Midwest populations had GxWeather, covarying with a cumulative rainfall and a photoperiod cue, respectively.

@@

keywords:  [allele-by-environment effect variation, antagonistic pleiotropy, photoperiod, cumulative rainfall, genetic variation]

@@

\firstpage[10]{2}

@@

Flowering time, in particular, is a common subject of GxE research [@blackman2013interacting; @henry2014transitions; @aagren2017adaptive; @brachi2010linkage; @dittmar2014flowering; @li2018genomic; @romero2017study], a key output of selection driving adaptation to local environments [@andres2012genetic; @wadgymar2017identifying; @blumel2015flowering], and a selection target for crop improvement to adapt crops to local or future environments @jung2009flowering. Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in a number of cereal species [@turner2005pseudo; @faure2012mutation; @hung2012zmcct; @zakhrabekova2012induced; @yang2013oself3] and other crops [@pin2012multifaceted; @weller2019parallel]. Recent statistical advances in studying phenological GxE have involved determining critical environmental indices before the phenological event occurs, such as photothermal time within a critical growth window @li2018genomic. However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE. Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying], but has been limited by a known statistical bias that reduced detection of genetic effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary]. Thus, despite substantial interest in the frequencies of various forms of GxE, the prevalence of antagonistic pleiotropy relative to other forms of GxE remains unknown.

@@

In our diversity panel of tetraploid switchgrass @lovell2021genomic, genotypes from the Gulf and Midwest genetic subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (@fig-map). At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, Gulf green-up occurred before Midwestern green-up, and Gulf flowering occurred after Midwestern flowering (@fig-map A). At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both Gulf green-up and flowering occurred after Midwest green-up and flowering. At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up occurred over the same time period. These patterns led to strong negative phenotypic correlations for green-up between the North and Texas gardens, particularly in the Gulf and Both subpopulations, and contributed to positive phenotypic correlations for flowering time of larger magnitude at more northern gardens (@fig-map B).

Narrow-sense heritabilities (h^2^) indicated that rank-changing GxE for these phenotypes was present across the common gardens (@fig-map C). h^2^ were typically high at individual gardens: 59% on average for green-up date, and 87% for flowering date. However, h^2^ were variable across gardens, and green-up dates were uncorrelated (r^2^ \< 0.2) or negatively correlated between pairs of gardens (@fig-map B). These negative and small correlations undoubtedly contributed to the low h^2^ values for green-up and flowering date when estimated jointly at all eight gardens: h^2^ was 0.8% for green-up and 23.2% for flowering date.

@@

We next explored how genetic variation in phenology covaried with environmental cues by using mash to jointly discover trait-associated SNPs across all common gardens for the Midwest, Gulf, and 'Both' subpopulations (@fig-map D, SI Appendix, Datasets 1-6). We first estimated SNP effects using GWAS on site-specific BLUPs for each phenotype. We then used mash on two subsets of these SNP effects, first to determine the covariance structures that significantly improved the mash model log likelihood when included, then to re-estimate the set of SNP effects with the largest effect estimates in univariate GWAS (hereafter "strong" effects) using the set of significant covariance matrices.

### 2023-12-28 manuscript/revision2/revision2.tex
manuscript: Edit Abstract and Significance statement

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Switchgrass Phenology and Genotype-by-Environment Mapping},
  pdfauthor={Alice H. MacQueen; Li Zhang; Samuel A. Smith; Jason Bonnette; Arvid R. Boe; Phillip A. Fay; Felix B. Fritschi; David B. Lowry; Robert B. Mitchell; Francis M. Rouquette Jr; Yanqi Wu; Arbel Harpak; Thomas E. Juenger},
  pdfkeywords={allele-by-environment effect variation, antagonistic
pleiotropy, photoperiod, cumulative growing degree days, genetic
variation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\affil[a]{University of Texas at Austin, Department of Integrative
Biology, Austin, 78712}
\affil[b]{South Dakota State University, Department of
Agronomy, Brookings}
\affil[c]{USDA-ARS, Grassland, Soil and Water Research
Laboratory, Temple}
\affil[d]{University of Missouri, Division of Plant Sciences, Columbia}
\affil[e]{Michigan State University, Department of Plant Biology, East
Lansing}
\affil[f]{USDA-ARS, Wheat, Sorghum, and Forage Research Unit, Lincoln}
\affil[g]{Texas A\&M University, Texas A\&M AgriLife Research and
Extension Center, Overton}
\affil[h]{Oklahoma State University, Department of Plant and Soil
Sciences, Stillwater}

@@

\significancestatement{The timing of plant seasonal development
(phenology) has major impacts on fitness because of the steep price of
plant-environment mismatches. Here we map the genetic basis of two
phenological events, the start of above-ground growth and flowering, in
two genetically and phenologically distinct populations of switchgrass.
We do this at eight field locations spanning the latitudinal range of
both switchgrass populations. Our approach allows us to identify regions
of the genome that covary with specific weather-related environmental
features at every location. For flowering, these features differed by
population: the Midwest population had genetic variation that primarily
covaried with cumulative growing degree days, a temperature-related
measure, while the Gulf population had genetic variation that primarily
covaried with photoperiod, a day-length-related measure.}

@@

\keywords{allele-by-environment effect variation | antagonistic
pleiotropy | photoperiod | cumulative growing degree days | genetic
variation}

@@

\begin{abstract}
The timing of vegetative growth and flowering in plants (``phenological
timings'') multiple depend on environmental cues; thus, phenological
traits can have important genotype-by-environment (GxE) interactions.
Here, we map genotype-by-weather effects on phenological timings in
multiple populations of two highly divergent switchgrass (Panicum
virgatum) populations spanning the central United States. We distinguish
GxWeather patterns covarying with interpretable weather-based cues and
agnostic, empirically-inferred GxWeather patterns. Most GxWeather
effects belong to the latter category and do not covary with
weather-based cues. However, \textgreater XX\% of Gulf population
effects on flowering time covary with photoperiod cues, while
\textgreater XX\% of Midwest upland population SNPs affecting flowering
had effects that covaried with XX cues. An independent pseudo-F2 cross
of Gulf and Midwest individuals mapped 23 additive QTLs for flowering at
the same common gardens, all with significant mash associations and ten
with enrichment of highly significant mash associations. We demonstrate
that we can identify QTL with GxE and assign them to specific
weather-based cues and/or data-driven patterns. Breeding for particular
alleles at these loci could change flowering responsiveness to
photoperiod cues in switchgrass. More broadly, this approach could be
used to identify genetic marker-environment interactions in any species
with related populations phenotyped in multiple environments.
\end{abstract}

@@

\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
\firstpage[15]{2}

@@

Flowering time, in particular, is a common subject of GxE research
(5--11), a key output of selection driving adaptation to local
environments (2, 12, 13), and a selection target for crop improvement to
adapt crops to local or future environments (14). Changing flowering
responsiveness to photoperiod cues has allowed geographic range
expansion and increased yields in a number of cereal species (15--19)
and other crops (20, 21). Recent statistical advances in studying
phenological GxE have involved determining critical environmental
indices before the phenological event occurs, such as photothermal time
within a critical growth window (10). However, most studies of flowering
GxE focus on finding a single, best fitting form of genotype-environment
covariance, despite the key expectation that different genetic
subpopulations, and even different genomic regions, have likely evolved
distinct patterns of GxE. Additionally, despite theoretical predictions
that local adaptation should involve antagonistic pleiotropy, or
sign-changing GxE, at the level of individual loci (22--25), previous
work has found limited evidence of antagonistic pleiotropy (12, 26), but
has been limited by a known statistical bias that reduced detection of
SNP effects that differ in sign (26--28). Thus, despite substantial
interest in the frequencies of various forms of GxE, the prevalence of
antagonistic pleiotropy relative to other forms of GxE remains unknown.

@@

In our diversity panel of tetraploid switchgrass (32), genotypes from
the Gulf and Midwest genetic subpopulations had distinct phenological
timings and distinct patterns of phenological correlations across our
eight common garden sites (Figure~\ref{fig-map}). At the three Texas
common gardens (hereafter `Texas' gardens), located within the natural
range of the Gulf subpopulation, Gulf green-up occurred before
Midwestern green-up, and Gulf flowering occurred after Midwestern
flowering (Fig. 1A). At the four northernmost common gardens (hereafter
`North' gardens), located within the natural range of the Midwest
subpopulation, both Gulf green-up and flowering occurred after Midwest
green-up and flowering. At the Oklahoma common garden, located near the
natural range limits of both the Gulf and the Midwest subpopulations,
Gulf and Midwest green-up occurred over the same time period. These
patterns led to strong negative phenotypic correlations for green-up
between the North and Texas gardens, particularly in the Gulf and Both
subpopulations, and contributed to positive phenotypic correlations for
flowering time of larger magnitude at more northern gardens (Fig. 1B).

Narrow-sense heritabilities (h2) indicated that rank-changing GxE for
these phenotypes was present across the common gardens (Fig 1C). h2 were
typically high at individual gardens: 59\% on average for green-up date,
and 87\% for flowering date. However, h2 were variable across gardens,
and green-up dates were uncorrelated (r2 \textless{} 0.2) or negatively
correlated between pairs of gardens (Fig. 1B). These negative and small
correlations undoubtedly contributed to the low h2 values for green-up
and flowering date when estimated jointly at all eight gardens: h2 was
0.8\% for green-up and 23.2\% for flowering date.

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Switchgrass Phenology and Genotype-by-Environment Mapping},
  pdfauthor={Alice H. MacQueen; Li Zhang; Samuel A. Smith; Jason Bonnette; Arvid R. Boe; Phillip A. Fay; Felix B. Fritschi; David B. Lowry; Robert B. Mitchell; Francis M. Rouquette Jr; Yanqi Wu; Arbel Harpak; Thomas E. Juenger},
  pdfkeywords={allele-by-environment effect variation, antagonistic
pleiotropy, photoperiod, cumulative rainfall, genetic variation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\affil[a]{University of Texas at Austin, Department of Integrative
Biology, Austin, 78712}
\affil[b]{South Dakota State University, Department of
Agronomy, Brookings, 57006}
\affil[c]{USDA-ARS, Grassland, Soil and Water Research
Laboratory, Temple, 76502}
\affil[d]{University of Missouri, Division of Plant
Sciences, Columbia, 65211}
\affil[e]{Michigan State University, Department of Plant Biology, East
Lansing, 48824}
\affil[f]{USDA-ARS, Wheat, Sorghum, and Forage Research
Unit, Lincoln, 68583}
\affil[g]{Texas A\&M University, Texas A\&M AgriLife Research and
Extension Center, Overton, 75684}
\affil[h]{Oklahoma State University, Department of Plant and Soil
Sciences, Stillwater, 74078}

@@

\significancestatement{The timing of plant seasonal development
(phenology) has major impacts on fitness because of the steep price of
plant-environment mismatches. We infer the mixture of ways genetic
effects on phenological traits depend on plant environment, particularly
how effects covary with weather prior to the phenological event
(GxWeather). Most effects do not have GxWeather, but a minority do.
GxWeather is population-specific. The majority of effects on the start
of vegetative growth in the Gulf population have GxWeather: these
effects differ in sign across environments and covary with a photoperiod
cue 14 days prior. A minority of effects on flowering date in the Gulf
and Midwest populations had GxWeather, covarying with a cumulative
rainfall and a photoperiod cue, respectively.}

@@

\keywords{allele-by-environment effect variation | antagonistic
pleiotropy | photoperiod | cumulative rainfall | genetic variation}

@@

\begin{abstract}
The timing of vegetative growth and flowering in plants (``phenological
timings'') depend on both genetic variation and environmental cues;
thus, phenological traits can have important genotype-by-environment
(GxE) interactions. In particular, understanding the mixture and
prevalence of genotype-by-weather (GxWeather) interactions primed by
weather prior to the phenological event should aid prediction and
manipulation of phenological timings. Here, we map genotype-by-weather
effects on phenological timings in two highly divergent switchgrass
(\emph{Panicum virgatum}) populations using repeated plantings of these
populations at sites spanning the central United States. We distinguish
GxWeather patterns covarying with interpretable weather-based cues from
agnostic, site-based patterns. Most GxWeather effects belong to the
latter category and do not covary with weather-based cues. However, 65\%
of Gulf population effects on the start of vegetative growth covary with
daylength 14 days prior to green-up date. 33\% of Gulf population
effects on flowering date covary with cumulative rainfall in the seven
days prior to flowering, while 22\% of Midwest population effects on
flowering date covaried with day length change two days prior to
flowering. An independent pseudo-F2 cross of Gulf and Midwest
individuals mapped 23 additive QTLs for flowering at the same common
gardens, all with significant mash associations and eleven with
enrichment of highly significant mash associations. We demonstrate that
we can identify QTL with GxWeather and assign them to specific
weather-based cues or other patterns. Breeding for particular alleles at
these loci could change flowering responsiveness to photoperiod cues in
switchgrass. More broadly, this approach could be used to identify
genetic marker-environment interactions in any species with related
populations phenotyped in multiple environments.
\end{abstract}

@@

\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
\firstpage[10]{2}

@@

Flowering time, in particular, is a common subject of GxE research
(5--11), a key output of selection driving adaptation to local
environments (2, 12, 13), and a selection target for crop improvement to
adapt crops to local or future environments (14). Changing flowering
responsiveness to photoperiod cues has allowed geographic range
expansion and increased yields in a number of cereal species (15--19)
and other crops (20, 21). Recent statistical advances in studying
phenological GxE have involved determining critical environmental
indices before the phenological event occurs, such as photothermal time
within a critical growth window (10). However, most studies of flowering
GxE focus on finding a single, best fitting form of genotype-environment
covariance, despite the key expectation that different genetic
subpopulations, and even different genomic regions, have likely evolved
distinct patterns of GxE. Additionally, despite theoretical predictions
that local adaptation should involve antagonistic pleiotropy, or
sign-changing GxE, at the level of individual loci (22--25), previous
work has found limited evidence of antagonistic pleiotropy (12, 26), but
has been limited by a known statistical bias that reduced detection of
genetic effects that differ in sign (26--28). Thus, despite substantial
interest in the frequencies of various forms of GxE, the prevalence of
antagonistic pleiotropy relative to other forms of GxE remains unknown.

@@

In our diversity panel of tetraploid switchgrass (32), genotypes from
the Gulf and Midwest genetic subpopulations had distinct phenological
timings and distinct patterns of phenological correlations across our
eight common garden sites (Figure~\ref{fig-map}). At the three Texas
common gardens (hereafter `Texas' gardens), located within the natural
range of the Gulf subpopulation, Gulf green-up occurred before
Midwestern green-up, and Gulf flowering occurred after Midwestern
flowering (Figure~\ref{fig-map} A). At the four northernmost common
gardens (hereafter `North' gardens), located within the natural range of
the Midwest subpopulation, both Gulf green-up and flowering occurred
after Midwest green-up and flowering. At the Oklahoma common garden,
located near the natural range limits of both the Gulf and the Midwest
subpopulations, Gulf and Midwest green-up occurred over the same time
period. These patterns led to strong negative phenotypic correlations
for green-up between the North and Texas gardens, particularly in the
Gulf and Both subpopulations, and contributed to positive phenotypic
correlations for flowering time of larger magnitude at more northern
gardens (Figure~\ref{fig-map} B).

Narrow-sense heritabilities (h\textsuperscript{2}) indicated that
rank-changing GxE for these phenotypes was present across the common
gardens (Fig 1C). h\textsuperscript{2} were typically high at individual
gardens: 59\% on average for green-up date, and 87\% for flowering date.
However, h2 were variable across gardens, and green-up dates were
uncorrelated (r2 \textless{} 0.2) or negatively correlated between pairs
of gardens (Fig. 1B). These negative and small correlations undoubtedly
contributed to the low h2 values for green-up and flowering date when
estimated jointly at all eight gardens: h2 was 0.8\% for green-up and
23.2\% for flowering date.
